---
title: "Assignment 3 - CRD"
author: "Jonathan S. Cardenas"  
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: false
    theme: cerulean
---

# Introduction

This assignment will ask you to conduct a full CRD analysis, from data import through publication ready plot.

The different sections representing what we did in class were left in the script to provide you a scaffold to start from.

# Instructions

-   Move this script into the `code` subfolder of your `03_crd` project folder\
-   Move the file `wheat_nk_bamyan.csv` into the `data` folder of your `03_crd` project folder\
-   On this script, fill your name under the `title` section of the YAML

# a) Setup

Here is where we load the packages we will use.\
Make sure to load each package in the chunk below.

```{r setup}
library(tidyverse) # for data wrangling and plotting
library(car) # for Anova function
library(broom) # for model residuals extraction
library(emmeans) # for model mean extraction
library(multcomp) # for pairwise comparison letter display

```

Now, import the data set `wheat_nk_bamyan.csv` in the chunk below.

```{r data import}
crd_df <- read_csv('../data/wheat_nk_bamyan.csv')

```

The **objective of this study** was to assess the effect of different N and K rates on wheat yield.

For that, a study was conducted using a completely randomized design as the experimental design with four replicates. The treatment design was a 3 N x 3 K rate factorial, for a total of 9 treatments.

# b) EDA tables

Use this space to conduct exploratory data analysis of the imported data, in the form of tables.

```{r}
summary(crd_df)
```

```{r}
glimpse(crd_df)
```

# c) Wrangling

Here, make sure to wrangle your data set with all needed steps. Refer back to class code if you need a refresher on what was needed.

```{r}
crd_dfw <- crd_df %>%
   mutate(rep = factor(rep),
         nrate_kgha = factor(nrate_kgha),
         krate_kgha = factor(krate_kgha)
         ) %>%
  mutate(trtname = paste0(nrate_kgha,'+',krate_kgha))
  
```

# d) EDA plots

Now, explore the wrangled data using plots.\
Make sure to explore plots showing the main effects and the interaction.

### Main effects

-   Nitrogen Rate\

```{r}
crd_dfw %>% ggplot(aes(nrate_kgha, yield_kgha)) +
  geom_boxplot() +
  geom_jitter(width = 0.2) + 
  theme_bw()
```

Color in jitter geom seems redundant butThere is not a significant in yield caused by increasing the nitrogen rate.

K rate

```{r}
crd_dfw %>% ggplot(aes(krate_kgha, yield_kgha)) +
  geom_boxplot() +
  geom_jitter(width = 0.2) + 
  theme_bw()
```

k rate has a more evident effect in yield. In fact, based on medians difference there is likely to be a significant difference between both 30 and 60 rates compared with no K.

### Interaction

```{r}
crd_dfw %>% ggplot(aes(nrate_kgha, yield_kgha, fill = krate_kgha)) +
  geom_boxplot() +
  geom_jitter(width = 0.3,alpha = 0.7) + 
  theme_bw()+
  facet_grid(.~krate_kgha) +
  scale_fill_brewer(palette = 2)
```

This faceted graph provides a clear hint that the interaction effect is significant. As we move right within the grid, the effect in yield seems to reach its maximum when 200 N rate is combined with either 30 or 60 kgha of K.

## Question 1) Below each plot of this section, write a sentence explaining your visual interpretation of the plot.

# e) Statistical model

In this section, run the appropriate model.\
Remember to set the contrast type to sum-to-zero (the default in R is set-to-zero, which is not what we want).

```{r}
options(contrast = c("contr.sum", "contr.poly"))

crd_mod <- lm(yield_kgha ~ nrate_kgha * krate_kgha,
              data = crd_dfw)

summary(crd_mod)
```

Also, remember to specify all model terms needed for the treatment design.

# f) ANOVA table

Use this section to check the ANOVA table, ensuring you use type 3 sum-of-squares.

```{r}
car::Anova(crd_mod, type = 3)
```

## Question 2) Based on the ANOVA table above, for which effect(s) in the ANOVA should you extract means and perform pair-wise comparisons? Explain why.

For the effects of K rate as this is the only significant effect among main and interaction effects.

# g) Linear model assumptions

Use this section to first compute your residuals, including studentdized residuals.

```{r}
crd_resid <- augment(crd_mod) %>%
  mutate(.studresid = rstudent(crd_mod))

crd_resid
```

## Question 3) What are the linear model assumptions that we need to check for? Create subsections (using ##) below for each of these assumptions, and write code to produce plots to check for these assumptions. For each assumption, write down your interpretation of the produced plots, and your thoughts if the assumptions hold or not.

## Residual independence

```{r}
crd_resid %>% ggplot(aes(x = .fitted,
                         y = .studresid)) +
  geom_point( shape = 21,
              fill = 'green4',
              size = 3,
              alpha = 0.6
              ) +
  geom_hline(yintercept = c(-3,0,3), color = 'red3') +
  geom_smooth() +
  theme_bw()
```

## Residual homoscedasticity

```{r}
crd_resid %>% ggplot(aes(x = .fitted,
                         y = .studresid)) +
  geom_point( shape = 21,
              fill = 'green4',
              size = 3,
              alpha = 0.6
              ) +
  geom_hline(yintercept = c(-3,0,3), color = 'red3') +
  geom_smooth() +
  theme_bw()
```

## Residual normality

```{r}
ggplot(crd_resid, aes(sample = .studresid)) +
  stat_qq(color = "red") +
  stat_qq_line()
```

```{r}
ggplot(crd_resid, aes(x = .studresid)) +
  geom_density(fill = "green4")+
  scale_x_continuous(breaks = c(-3,0,3), limits = c(-3,3))
```

## Residual Outliers

# h) Model means

Use this section to extract model means according to your decision of the ANOVA table effects.

If we recall the study objective: assess the effect of different N and K rates on wheat yield. Based on the ANOVA we can conclude that the effect of interactions or the effects of N rates are not significant on wheat yield. The only significant effect are the different K rates. 

```{r}
crd_means_k <- emmeans(crd_mod, ~ krate_kgha)
crd_means_k
```


# i) Pairwise comparisons

Use this section to perform pair-wise comparison of the means extracted above.

Make sure to reverse the order from high to low, use letters instead of numbers, and no adjustment.

```{r}
crd_cld_k <- cld(crd_means_k, 
                 reversed = T,
                 Letters = letters,
                 adjust = 'none')
crd_cld_k
```


# g) Final plot(s)

Produce final plot(s) to show your results.\
Make sure to include both **raw data** and **statistical model summary (i.e., letter separation)**.

Be mindful of your plot to ensure you are creating it in an intuitive way for your audience in relation to interpreting the letter separation structure you used.

In class, we did not have time to make it publication-ready (e.g., a plot looking professional and ready to be included in a manuscript).

Here, make sure you deliver final plot(s) that is/are publication ready.

Make sure to provide an interpretation for the plot(s) created, including which treatment levels were significantly higher/lower.


```{r}
ggplot(mapping = aes(fill = krate_kgha )) +
  geom_boxplot(data = crd_dfw,
               aes(x = krate_kgha, 
                   y = yield_kgha ),
               alpha = 0.6
               ) +
  geom_jitter(data = crd_dfw,
              aes(x = krate_kgha,
                  y = yield_kgha),
              shape = 21,
              size = 2,
              alpha = 0.8,
              width = 0.2
              ) +
  geom_label(data= crd_cld_k,
             aes(x = krate_kgha,
                 y = emmean,
                 label = .group),
             fill = "white"
             ) +
  theme_bw() +
  theme(legend.position = "null") +
   labs(x = "K rate (kg/ha)",
       y = "Yield (kg/ha)"
       ) +
  scale_fill_viridis_d(option = "E")
  
```
As K rate treatment factor was the only significant effect,  levels indicate a rate of 30 and 60 kg/ha increase wheat yield significantly compared with no K fertilization.  

# Submitting your work

Once you have developed all the code and answers, make sure to Render this quarto file.

**Notes on rendering**:

-   Make sure to render your work and inspect how the final html look like.\
-   If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.\
-   Only send me your work once your html file looks professional.
    -   **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.\

    -   If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).

    -   If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).

After rendering, an .html file will be created on your `code` folder.

Rename this file to `Assignment-03-crd-LASTNAME.html`.

For ex., mine would be `Assignment-03-crd-Bastos.html`.

Send the **html** file to **lmbastos\@uga.edu** by Feb 15th 11:59 pm.
